---
title: "reproducible_papers_with_rmarkdown"
author: "Apryl Perry"
date: "August 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If pkg is not avaialbe on CRAN b/c it is still in developement, you can still get the developement version from github using the following:

```{r}
install.packages("captioner") #for storing figure and table captions
devtools::install_github("benmarwick/rrtools") #devlopement version not available as a CRAN package yet
rrtools::create_compendium("testpaper") #creates a new R pkg with a directory structure ready to perform analysis
```

Next go to new RMarkdown that has opened and navigate to analysis folder, then paper, open paper.rmd file to modify as desired.  Note: data should be stored in an archival repository and should be access from there, not from the data folder included in this file directory!

Create empty repository on github, do not initalize
In R terminal window run
git remote add origin and paste the url for the repository in
git remote fetch origin

