---
title: "Tidying Data in R"
author: "Apryl Perry"
date: "August 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

can delete other info that comes at the start at rmarkdown, but need to leave the above setup code

Calling a masked function:

If you call filter w/o calling stats, b/c of the way the pkgs were loaded.  These functions do specifically different things.  See notes in yellow above (currently suppressed by the command options warning and message and echo above).

use
packageName::functionName
or 
stats::filter()
depending on which package function you want to use.

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

Read in data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2] (https://knb.ecoinformatics.org/#view/df35b.304.2).

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

Data in the above output is not Tidy.  Fish names in columns and All column is a sum column


About the pipe editor
The pipe will take the output of a previous statement, and use it as the input to the next statement. 
ctrl-shift-m is keyboard shortcut for pipe operator
chains operations or functions together
ctrl dash shortcut for assignment operator

df_filtered <- filter(df, ...)
df_selected <- select(df_filtered, ...)

can write:
df_cleaned <- df %>% 
              filter(...) %>%
              select(...)
              
In the Tidy verse, all data is assumed to be a data frame (that is what the (.data, ...)) part of the code represents, but you will probably never explicitly state this.
Select the columns you want by name              
```{r}
catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

```

or specify the column you don't want by using the - sign.
```{r}

catch_df %>% 
  select(-All, -notesRegCode)
  
head(catch_df)

```
Need to place %>%  at end of the line if adding another function.  Place the next function underneath for 
easier readability.

```{r}
catch_df <- catch_df %>% 
  select(-All, -notesRegCode) # %>% 
  # filter(...)
head(catch_df)

```

key = column name, value = value in columns
What am i turning the column names into?
Gather will only return two columns unless you specifically tell it otherwise, -Region, -Year will cause the info from these columns to also be displayed. - in the gather statement means don't write over these columns.
rename function without using index or list selector operator
rename catch column to include units
```{r}
catch_df <- catch_df %>% 
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousand = catch)

unique(catch_df$species) # shows the names of unique names in species column
head(catch_df)
``` 

reverse of gather is spread function

note that the catch_thousands is a character column and not numeric so mutate returns an error.  

```{r}
catch_integers <-  as.integer(catch_df$catch_thousand)

i <- which(is.na(catch_integers) == T)

catch_df$catch_thousand[i]
```

Warning above is due to not being able to turn some values into a number.
second line of code allows you to inspect what happened above and only returns the values that are true to the above statement
third line allows you to inspect those items

The mutate function allows you to create new columns but to fix above issue, need to:
1) need to turn I into 1 in the catch_thousand column
2) coerce catch_thousand into integer type
3) convert catch_thousand to catch

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousand = ifelse(catch_thousand == "I", 1, catch_thousand)) %>% 
  mutate(catch_thousand = as.integer(catch_thousand)) %>% 
  mutate(catch = catch_thousand * 1000)

tail(catch_df)
# head(catch_df)
# summary(catch_df)
```


alt-shift-k gives a list of keyboard shortcuts in RStudio
ctrl-shift-= zooms all panes (seems to do all of computer also...)

#Split_Apply_Combine
Performing mean function on catch column by region
Only has two columns b/c we have averaged over the year and species column
with summarise, the output needs to be one value (note will take both english and british spelling of summarize)

```{r}
catch_sum <- catch_df %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))
head(catch_sum, 10) #Shows the first 10 rows instead of the default setting
```
Create catch low and catch high columns averaged over species and years for both columns

```{r}
catch_range <- catch_df %>% 
  group_by(Region) %>% 
  summarise(catch_low = range(catch)[1],
            catch_high = range(catch)[2])
head(catch_range, 10)
```

or grouped by year

```{r}
catch_range <- catch_df %>% 
  group_by(Year) %>% 
  summarise(catch_low = range(catch)[1],
            catch_high = range(catch)[2])
head(catch_range, 10)
```
What is the number of observations in each grouping

```{r}
catch_mean <- catch_df %>% 
  group_by(Region) %>% 
  #calculate mean, sd, n
  summarise(catch_mean = mean(catch, na.rm = T),
            catch_stdev = sd(catch, na.rm = T),
            num_obs = n())
head(catch_mean, 10)
```

```{r}
catch_mean <- catch_df %>% 
  group_by(Region, species) %>% #Group by Region and species
  #calculate mean, sd, n
  summarise(catch_mean = mean(catch, na.rm = T),
            catch_stdev = sd(catch, na.rm = T),
            num_obs = n())
head(catch_mean, 10)
```
Filter

```{r}
catch_mean <- catch_df %>% 
  group_by(Region, species) %>% 
  summarise(catch_mean = mean(catch, na.rm = T),
            catch_stdev = sd(catch, na.rm = T),
            num_obs = n()) %>% 
#filter for Chinook
filter(species == "Chinook") %>% 
  filter(catch_mean > 2000)
head(catch_mean, 10)

```
```{r}
catch_year <-  catch_df %>% 
  filter(Year > 1990 | Year < 1900) %>% # use & for and currently is filtering for > 1990 or < 1900
  group_by(Year, species) %>% 
  summarize(catch_sum = sum(catch))

head(catch_year)
# print(catch_year) shows all records
```

Joins
Left join will return all of the rows from the left side table and add on any columns from the right side using the table key (in this case the key is the Region column)
If there is no matching value in the right side, NA will be filled in instead as it maintains data set integrity.  This is also a good way to find potential issues in the data set. 
Joins should be extremely predictable
If left join added rows, then something went wrong.  If you have dupliate identifiers in the identifyer row, you will get duplicate rows...potentially a lot of them

If there are more rows in the right table, they will be discarded (with a left join)

```{r} 
# Read in data from URL
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```

only interested in management column use select function
```{r}
region_defs <-  region_defs %>% 
  select(code, mgmtArea)

head(region_defs)
```

```{r}
#If both columns being used in the join have the exact same name, R can guess, otherwise they need to be specified as shown below.  Can have multiple key columns, just separate by a comma
catch_joined <- left_join(catch_df, region_defs, by = c("Region" = "code"))
    
head(catch_joined)
```












