---
title: "Tidying Data in R"
author: "Apryl Perry"
date: "August 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

can delete other info that comes at the start at rmarkdown, but need to leave the above setup code

Calling a masked function:

If you call filter w/o calling stats, b/c of the way the pkgs were loaded.  These functions do specifically different things.  See notes in yellow above (currently supressed by the command options warning and message and echo above).

use
packageName::functionName
or 
stats::filter()
depending on which package function you want to use.

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

Read in data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2] (https://knb.ecoinformatics.org/#view/df35b.304.2).

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```



